---
title: "invaderCheck-Lite"
author: "Sunit"
date: "04/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(scales)
library(lubridate)
```


```{r}
workdir = "/Users/sunit.jain/Research/Alice/InvaderCheckLite"
fig_dir = paste0(workdir,"/Figures")

min_reads = 2
min_abundance = 1e-4
time_stamp = now() %>% format("%Y%m%d")
mbfv1_midas_output = "/Users/sunit.jain/Research/Alice/in_vivo/Mouse_Backfill/Midas_Results/v1/20201210_mbfv1_dataframe_minRelAbund_0.combined.csv"
mbfv2_midas_output = "/Users/sunit.jain/Research/Alice/in_vivo/Mouse_Backfill/Midas_Results/v2/20201204_mbfv2_dataframe_minRelAbund_0.csv"

scv1_2.strain_contributions = "/Users/sunit.jain/Research/Sunit/Midas/SCv1_2.art_reads_output.full_dataframe.csv"
scv2_3.strain_contributions = "/Users/sunit.jain/Research/Sunit/Midas/SCv2_3.art_reads_output.full_dataframe.csv"
# mbfv1_p100 = "/Users/sunit.jain/Research/Alice/Midas_NinjaMap_ImmiCheck_Compare/01_vectors_coverage_stats.csv"
mbfv1_p100 = "/Users/sunit.jain/Research/Alice/InvaderCheckLite/MBFv1/genome_coverage_summary.id100_aln100.csv"
mbfv1_p99 = "/Users/sunit.jain/Research/Alice/InvaderCheckLite/MBFv1/genome_coverage_summary.id99_aln100.csv"
mbfv2_p100 = "/Users/sunit.jain/Research/Alice/InvaderCheckLite/MBFv2/genome_coverage_summary.id100_aln100.csv"
mbfv2_p99 = "/Users/sunit.jain/Research/Alice/InvaderCheckLite/MBFv2/genome_coverage_summary.id99_aln100.csv"
```


```{r}
make_dir = function(dir_path){
  if (! dir.exists(dir_path)){
    dir.create(dir_path, recursive = TRUE, mode = "0777")
  }
  return(dir_path)
}

make_dir(fig_dir)
```


```{r}
new_genomes_in_SCv2_3 = c(
"Adlercreutzia-equolifaciens-DSM-19450",
"Alistipes-finegoldii-DSM-17242",
"Alistipes-ihumii-AP11-MAF-2",
"Alistipes-indistinctus-YIT-12060-DSM-22520-MAF-2",
"Alistipes-onderdonkii-DSM-19147-MAF-2",
"Alistipes-senegalensis-JC50-DSM-25460-MAF-2",
"Alistipes-shahii-WAL-8301-DSM-19121-MAF-2",
"Bacteroides-rodentium-DSM-26882-MAF-2",
"Bilophila-wadsworthia-ATCC-49260",
"Blautia-sp-KLE-1732-HM-1032-MAF-2",
"Blautia-wexlerae-DSM-19850-MAF-2",
"Bryantella-formatexigens-DSM-14469-MAF-2",
"Burkholderiales-bacterium-1-1-47",
"Butyricimonas-virosa-DSM-23226-MAF-2",
"Clostridiales-bacterium-VE202-03",
"Clostridiales-bacterium-VE202-14",
"Clostridiales-bacterium-VE202-27-MAF-1",
"Clostridium-asparagiforme-DSM-15981-MAF-2",
"Clostridium-bartlettii-DSM-16795",
"Clostridium-bolteae-ATCC-BAA-613-MAF-2",
"Clostridium-sp-ATCC-29733-VPI-C48-50-MAF-1",
"Eubacterium-eligens-ATCC-27750-MAF-2",
"Intestinimonas-butyriciproducens-DSM-26588-MAF-2",
"Odoribacter-splanchnicus-DSM-20712",
"Oscillibacter-sp-KLE-1728-MAF-1",
"Ruminococcus-gauvreauii-DSM-19829-MAF-2",
"Ruminococcus-obeum-ATCC-29174-MAF-1",
"Subdoligranulum-sp-4-3-54A2FAA-MAF-1"
)
```


```{r}
read_midas_output = function(midas_output_file, minimum_reads, minimum_abundance, selected_week=NULL){
  df = read_csv(midas_output_file) %>% 
    filter(!is.na(Challenge)) %>% 
    filter(count_reads > minimum_reads) %>% 
    filter(relative_abundance > minimum_abundance) %>% 
    rowwise() %>% 
    mutate(sample_name = nth(str_split(sample_id,"_")[[1]], 1)) %>% 
    ungroup() %>% 
    select(-sample_id) %>% 
    rename("midas_bucket"="name",
           "sample_id"="sample_name")
    
  
  if (! is_null(selected_week)) {
    return (df %>% filter(Week==selected_week))
  } else {
    return (df)
  }
}

# read_midas_output(mbfv1_midas_output,min_reads, min_abundance, selected_week="W8")
```


```{r}
read_strain_contributions = function(strain_contributions_file){
  strains_df = read_csv(strain_contributions_file) %>%
  rename("midas_bucket"="name", 
         "genome_name" = "sample_id")
  
  return (strains_df)
}
```


```{r}
read_strain_contributions(scv2_3.strain_contributions) %>%
  group_by(genome_name) %>% 
  summarise(num_midas_buckets = n(),
            gross_contributions = sum(abundance, na.rm=TRUE)) %>% 
  arrange(-gross_contributions) %>% 
  filter(num_midas_buckets > 1)
# bucket_contributors
```
```{r}
read_strain_contributions(scv2_3.strain_contributions) %>%
  group_by(midas_bucket) %>% 
  summarise(num_strains = n(),
            gross_contributions = sum(abundance, na.rm=TRUE)) %>% 
  arrange(-gross_contributions) %>% 
  filter(num_strains > 1)
```

```{r}
extend_midas_output = function(midas_output_file, strain_contributions_file, minimum_reads, minimum_abundance, selected_week=NULL){
  midas_df = read_midas_output(midas_output_file, minimum_reads, minimum_abundance, selected_week)
  strains_midas_df = read_strain_contributions(strain_contributions_file)
  
  left_join(midas_df, strains_midas_df, by="midas_bucket") %>% 
    drop_na(genome_name)
}
# extend_midas_output(mbfv1_midas_output,scv1_2.strain_contributions, min_reads, min_abundance, selected_week="W8")
```


```{r}
read_coverage_stats = function(coverage_stats_file){
  read_csv(coverage_stats_file) %>% 
    rename("sample_id"="sample_name")
}
# read_coverage_stats(mbfv1_p100)
```

## IC-Lite Algorithm

- Compute the actual Midas results for each Week 8 sample [Midas --> Sample Details] -- (A)
- Compute Midas bucket contributions based on in silico reads. [Midas --> Strains] -- (B)
- Merge A and B on Midas Buckets [Sample Details <-- Midas --> Strains] -- (C)

- Compute the % coverage and depth of cov for each genome in each Sample [Strains --> Strain Details] -- (D)
- Merge C and D on Strains and Sample ids [Sample Details <-- Midas --> Strains --> Strain Details] -- (E)

- Plot (E)

```{r}
run_icLite_aggregation = function(midas_output_file, strain_contributions_file, coverage_stats_file, minimum_reads, minimum_abundance, selected_week){
  left_join(
    extend_midas_output(midas_output_file, strain_contributions_file, minimum_reads, minimum_abundance, selected_week),
    read_coverage_stats(coverage_stats_file),
    by = c("genome_name", "sample_id")
  )
}
```

```{r}
plot_icl_aggregation = function(df, title, subtitle=NULL){
  df %>% 
    ggplot(aes(ref_percent_coverage, mean_ref_coverage_depth))+
    geom_point(aes(color=relative_abundance), alpha = 0.3)+
    # scale_y_log10()+
    geom_rug(alpha=0.1)+
    scale_color_viridis_c(name="Midas \nrelative \nabundance")+
    scale_y_continuous(trans='log10', breaks = 10^(seq(-10,10,1)),
                       labels = trans_format("log10", math_format(10^.x)))+
    xlab("Percent genome covered")+
    ylab("Mean coverage depth")+
    ggtitle(title, subtitle)+
    theme_bw()+
    theme(text = element_text(size = 15),
          axis.text.x=element_text(size = 12),
          axis.text.y=element_text(size = 12))
}

```


### hCom1

```{r}
run_icLite_aggregation(mbfv1_midas_output,scv1_2.strain_contributions, mbfv1_p100, min_reads, min_abundance, selected_week="W8") %>%
  select(midas_bucket, genome_name, sample_id, relative_abundance, ref_percent_coverage, mean_ref_coverage_depth) %>%
  filter((relative_abundance >= min_abundance)) %>%
  plot_icl_aggregation(.,"hCom1 genomes in MBFv1 - Week8", "p100, aln100")

ggsave(paste0(fig_dir,"/",time_stamp,"_hCom1_mbfv1_w8.ic_lite_p100_a100.eps"),
       dpi="retina",
       height = 12,
       width = 16,
       units = "in")

ggsave(paste0(fig_dir,"/",time_stamp,"_hCom1_mbfv1_w8.ic_lite_p100_a100.pdf"),
       useDingbats=FALSE,
       dpi="retina",
       height = 12,
       width = 16,
       units = "in")
```

```{r}
run_icLite_aggregation(mbfv1_midas_output,scv1_2.strain_contributions, mbfv1_p99, min_reads, min_abundance, selected_week="W8") %>% 
  filter((relative_abundance >= min_abundance)) %>%
  plot_icl_aggregation(.,"hCom1 genomes in MBFv1 - Week8", "p99, aln100")

ggsave(paste0(fig_dir,"/",time_stamp,"_hCom1_mbfv1_w8.ic_lite_p99_a100.eps"),
       dpi="retina",
       height = 12,
       width = 16,
       units = "in")

ggsave(paste0(fig_dir,"/",time_stamp,"_hCom1_mbfv1_w8.ic_lite_p99_a100.pdf"),
       useDingbats=FALSE,
       dpi="retina",
       height = 12,
       width = 16,
       units = "in")
```
### hCom2

```{r}
run_icLite_aggregation(mbfv2_midas_output,scv2_3.strain_contributions, mbfv2_p100, min_reads, min_abundance, selected_week="Week8") %>% 
  filter((relative_abundance >= min_abundance)) %>%
  plot_icl_aggregation(.,"hCom2 genomes in MBFv2 - Week8", "p100, aln100")

ggsave(paste0(fig_dir,"/",time_stamp,"_hCom2_mbfv2_w8.ic_lite_p100_a100.eps"),
       dpi="retina",
       height = 12,
       width = 16,
       units = "in")

ggsave(paste0(fig_dir,"/",time_stamp,"_hCom2_mbfv2_w8.ic_lite_p100_a100.pdf"),
       useDingbats=FALSE,
       dpi="retina",
       height = 12,
       width = 16,
       units = "in")
```

```{r}
run_icLite_aggregation(mbfv2_midas_output,scv2_3.strain_contributions, mbfv2_p99, min_reads, min_abundance, selected_week="Week8") %>% 
  filter((relative_abundance >= min_abundance)) %>%
  plot_icl_aggregation(.,"hCom2 genomes in MBFv2 - Week8", "p99, aln100")

ggsave(paste0(fig_dir,"/",time_stamp,"_hCom2_mbfv2_w8.ic_lite_p99_a100.eps"),
       dpi="retina",
       height = 12,
       width = 16,
       units = "in")

ggsave(paste0(fig_dir,"/",time_stamp,"_hCom2_mbfv2_w8.ic_lite_p99_a100.pdf"),
       useDingbats=FALSE,
       dpi="retina",
       height = 12,
       width = 16,
       units = "in")
```

### (28?) new genomes in hCom2

```{r}
run_icLite_aggregation(mbfv2_midas_output,scv2_3.strain_contributions, mbfv2_p100, min_reads, min_abundance, selected_week="Week8") %>% 
  filter((relative_abundance >= min_abundance)) %>%
  filter(genome_name %in% new_genomes_in_SCv2_3) %>% 
  plot_icl_aggregation(.,"new hCom2 genomes in MBFv2 - Week8", "p100, aln100")

ggsave(paste0(fig_dir,"/",time_stamp,"_hCom2-new-genomes_mbfv2_w8.ic_lite_p100_a100.eps"),
       dpi="retina",
       height = 12,
       width = 16,
       units = "in")

ggsave(paste0(fig_dir,"/",time_stamp,"_hCom2-new-genomes_mbfv2_w8.ic_lite_p100_a100.pdf"),
       useDingbats=FALSE,
       dpi="retina",
       height = 12,
       width = 16,
       units = "in")
```


```{r}
run_icLite_aggregation(mbfv2_midas_output,scv2_3.strain_contributions, mbfv2_p99, min_reads, min_abundance, selected_week="Week8") %>%
  filter((relative_abundance >= min_abundance)) %>%
  filter(genome_name %in% new_genomes_in_SCv2_3) %>% 
  plot_icl_aggregation(.,"new hCom2 genomes in MBFv2 - Week8", "p99, aln100")

ggsave(paste0(fig_dir,"/",time_stamp,"_hCom2-new-genomes_mbfv2_w8.ic_lite_p99_a100.eps"),
       dpi="retina",
       height = 12,
       width = 16,
       units = "in")

ggsave(paste0(fig_dir,"/",time_stamp,"_hCom2-new-genomes_mbfv2_w8.ic_lite_p99_a100.pdf"),
       useDingbats=FALSE,
       dpi="retina",
       height = 12,
       width = 16,
       units = "in")
```

